---
title: "p8105_hw3_ew2717"
author: "Eric Wang"
output: github_document
---

```{r setup, include = FALSE}

knitr::opts_chunk$set(echo = TRUE)

```


```{r setup two}

library(tidyverse)
library(p8105.datasets)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = 0.6,
  out.width = "100%"
)

theme_set(theme_bw() + theme(legend.position = "bottom"))
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_color_viridis_d
scale_fill_discrete = scale_fill_viridis_d

```

## Problem 1

#### Loading instacart data

```{r}

data("instacart")

instacart_df <-
  instacart %>% 
  janitor::clean_names()

```

#### Instacart data exploration 

###### Finding rows, columns, and key variables.

```{r intstacart rows cols names}

instacart_rows = instacart_df %>% 
  nrow()
instacart_cols = instacart_df %>% 
  ncol()
instacart_names = instacart_df %>% 
  names()

```

The number of rows in the instacart data set is **`r instacart_rows`**. The number of columns in the instacart data set is **`r instacart_cols`**. The key variables in the instacart data set are **`r instacart_names`**.

###### Finding how many aisles there are and which aisles are the most items ordered from

```{r instacart aisles}

instacart_aisles = 
  instacart_df %>% 
  group_by(aisle) %>% 
  summarize(n_obs = n()) %>% 
  arrange(desc(n_obs))

instacart_aisles

```

As seen above, there are 134 aisles, and the aisles that have the most items ordered from are the "fresh vegetables", "fresh fruits", "packaged vegetable fruits", "yogurt", and "packaged cheese". 


###### Creating a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”, including the number of times each item is ordered

```{r aisle table}

aisle_table = 
  instacart_df %>% 
  group_by(aisle) %>% 
  summarize(item_count = n()) %>% 
  filter(item_count > 10000) %>% 
  ggplot(aes(x = reorder(aisle, item_count), y = item_count)) +
  geom_bar(stat = "identity") +
  labs(
    title = "Items ordered per Aisle",
    x = "Aisle",
    y = "# of Items",
    caption = "Figure 1. The table depicts the number of items ordered per aisle (>10000 items)."
  ) +
  coord_flip() +
  theme(axis.text.y = element_text(size = 6))

aisle_table

```

###### Making a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits", including the number of times each item is ordered


### Table of three most popular items in each of the aisles "baking ingredients", "dog food care", "packaged vegetables fruits"

```{r popular_items_table, message = FALSE, warning = FALSE}

popular_table = 
  instacart_df %>% 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>% 
  group_by(aisle, product_name) %>% 
  summarize(item_count = n()) %>% 
  filter(min_rank(desc(item_count)) < 4) %>% 
  arrange(desc(item_count)) 
  
knitr::kable(popular_table)

```

## PROBLEM 2 

#### Loading BRFSS dataset from p8105.datasets 

```{r load brfss data}

data("brfss_smart2010")

```

#### Cleaning brfss data set 

```{r clean brfss data set}

brfss_df = brfss_smart2010 %>% 
  janitor::clean_names() %>% 
  filter(
  topic == "Overall Health",
  response %in% c("Poor", "Fair", "Good", "Very good", "Excellent")) %>% 
  mutate(response = as.factor(response), response = ordered(response, levels = c("Poor", "Fair", "Good", "Very good", "Excellent"))) %>% 
  select(year, locationdesc, topic, response, data_value, sample_size) %>%
  separate(locationdesc, into = c("State", "County"), sep = " - ")

```

###### Finding which states were observed at 7 or more locations in 2002 and 2010

```{r observed 7 or more}

brfss_df_02 = 
  brfss_df %>% 
  group_by(State) %>% 
  filter(year == "2002") %>%
  distinct(County) %>%
  count(State) %>% 
  filter(n >= 7) %>% 
  arrange(n)

brfss_df_10 = 
  brfss_df %>% 
  group_by(State) %>% 
  filter(year == "2010") %>%
  distinct(County) %>%
  count(State) %>% 
  filter(n >= 7) %>% 
  arrange(n)

knitr::kable(brfss_df_02)
knitr::kable(brfss_df_10)

```

As the tables depict, CT, FL, NC, MA, NJ, and PA were observed at 7 or more locations in 2002, and CA, CO, FL, MA, MD, NC, NE, NJ, NY, OH, PA, SC, TX, and WA were observed at 7 or more locations in 2010.

###### Constructing a dataset that is limited to Excellent responses, and contains, year, state, and a variable that averages the data_value across locations within a state. 

```{r excellent dataset}

brfss_excellent_df = 
  brfss_df %>%
  filter(response == "Excellent") %>% 
  group_by(State, year) %>% 
  mutate(avg_data_value = mean(data_value, na.rm = TRUE),
         avg_data_value = round(avg_data_value, digits = 4)) %>%
  select(year, State, avg_data_value) 

brfss_excellent_df
```


